---
output:
  html_document:
    df_print: paged
    fig.align: center
    self_contained: yes 
    fig.height: 4
    fig.width: 8
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
    code_folding: hide
title: "Reproducibility rev001"
author: "David Lilek"
date: "`r format(Sys.time(), '%d %B %Y, %X')`"
editor_options: 
  markdown: 
    wrap: 72
---

# Load Libraries & Plot function & Colors

```{r message=FALSE, warning=FALSE}
# load libaries
library(GGally)
library(corrplot)
library(ggplot2)
library(pheatmap)
library(tidyverse)
library(dplyr)
library(UpSetR)
library(knitr)
library(reshape2)
library(eulerr)
library(UniprotR)
library(stringr)
library(ggfittext)
library(plotly)
library(pcaMethods)
```

# l-540 living

## Control

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(320:334)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```



## Etoposide

```{r}
############################################################
# E
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]


# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(335:349)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```



## RE

```{r}
############################################################
# RE
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(365:379)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```


## R

```{r}
############################################################
# R
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]


# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(350:364)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Compare All

```{r}
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

df.prot[df.prot == 0] <- NA

df.LFQ <- df.prot[,c(320:379)]
```



### CV calculation

```{r}
n <- ncol(df.LFQ)   
block_sizes <- rep(c(9, 6), length.out = ceiling(n / 6))
block_sizes <- block_sizes[cumsum(block_sizes) <= n]
if (length(block_sizes) == 0) {
  stop("too less columns")
}

ends <- cumsum(block_sizes)
starts <- c(1, head(ends, -1) + 1)

# generate df mean
df_mean <- as.data.frame(
  sapply(seq_along(starts),
         function(i) rowMeans(df.LFQ[, starts[i]:ends[i], drop = FALSE], na.rm = TRUE))
)

# col names
colnames(df_mean) <- paste0("block", seq_along(starts))

# calculate cv per biological replicate
cv_fun <- function(x) {
  # falls mean == 0 -> NA vermeiden
  m <- mean(x, na.rm = TRUE)
  if (is.na(m) || m == 0) return(NA_real_)
  sd(x, na.rm = TRUE) / m
}
# define pairs
pairs <- list(c(1,2), c(3,4), c(5,6), c(7,8))
# check if columns exist
pairs <- Filter(function(cols) all(cols <= ncol(df_mean)), pairs)

# rowwise cv
cv_rowwise <- sapply(pairs, function(cols) {
  apply(df_mean[, cols, drop = FALSE], 1, function(x) cv_fun(x))
})
cv_rowwise <- as.data.frame(cv_rowwise)
# name columns
colnames(cv_rowwise) <- c("Control", "Etoposide", "Resveratrol", "RE")[seq_along(pairs)]

# long format for plotting
cv_long <- pivot_longer(as.data.frame(cv_rowwise),
                        cols = everything(),
                        names_to = "Treatment",
                        values_to = "cv")

ggplot(cv_long, aes(x = Treatment, y = cv)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    x = "Spaltenpaar",
    y = "CV"
    )


colMeans(cv_rowwise*100, na.rm = TRUE)
apply(cv_rowwise*100, 2, median, na.rm = TRUE)

result_list <- lapply(colnames(cv_rowwise), function(col) {
  
  cv_values <- cv_rowwise[[col]]
  
  data.frame(
    Group = col,
    Median_CV = median(cv_values, na.rm = TRUE),
    IQR_CV = IQR(cv_values, na.rm = TRUE),
    Range_CV = max(cv_values, na.rm = TRUE) - min(cv_values, na.rm = TRUE),
    CV_lt_10 = mean(cv_values < 0.1, na.rm = TRUE) * 100,
    CV_lt_20 = mean(cv_values < 0.2, na.rm = TRUE) * 100,
    CV_gt_30 = mean(cv_values > 0.3, na.rm = TRUE) * 100
  )
})

summary_df <- do.call(rbind, result_list)
summary_df
```

### PCA 


```{r}
df.LFQ <- df.prot[,c(320:379)]
df.LFQ <- log2(df.LFQ)
fn <- colnames(df.LFQ)

# shorten filenames
fn_short <- sub("^LFQ\\.intensity\\.", "", fn)
fn_short <- sub("^.*LE_", "", fn_short)
fn_short <- sub("\\.raw\\.thermo\\.raw$", "", fn_short)

# rename them
colnames(df.LFQ) <- fn_short


mat <- df.LFQ[,c(1:9,
                 16:24,
                 31:39,
                 46:54)]
colnames(mat)

mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
#mat[mat == "NaN"] <- NA
#mat <- llsImpute(as.matrix(mat), k = 5, correlation = "pearson", allVariables = TRUE)
#mat <- as.data.frame(completeObs(mat))
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)


plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",      
  text = ~Filename,
  textposition = "top center",  
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```

```{r}
mat <- df.LFQ[,-c(1:9,
                 16:24,
                 31:39,
                 46:54)]

colnames(mat)

mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
#mat[mat == "NaN"] <- NA
#mat <- llsImpute(as.matrix(mat), k = 5, correlation = "pearson", allVariables = TRUE)
#mat <- as.data.frame(completeObs(mat))
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)


plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",      
  text = ~Filename,
  textposition = "top center",  
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```


# l-540 cryo

## Control

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]


# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(260:262, 266:271)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:3
B <- 4:9

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Etoposide

```{r}
############################################################
# E
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(272:277, 278:283)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:7
B <- 8:12

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## RE

```{r}
############################################################
# RE
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(296:301, 302:307)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:6
B <- 7:12

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## R

```{r}
############################################################
# R
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(284:289, 290:295)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:6
B <- 7:12

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Compare All

```{r}
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

df.prot[df.prot == 0] <- NA

df.LFQ <- df.prot[,c(260:262, 266:271,
                     272:277, 278:283,
                     296:301, 302:307,
                     284:289, 290:295)]
```

### CV caluclation

```{r}
n <- ncol(df.LFQ)

block_sizes <- c(3, 6, 6, 6, 6, 6, 6, 6)

if (sum(block_sizes) != n) stop("block sizes do not match column number")

ends <- cumsum(block_sizes)
starts <- c(1, head(ends, -1) + 1)

starts
ends

# generate df mean
df_mean <- as.data.frame(
  sapply(seq_along(starts),
         function(i) rowMeans(df.LFQ[, starts[i]:ends[i], drop = FALSE], na.rm = TRUE))
)

# col names
colnames(df_mean) <- paste0("block", seq_along(starts))

# calculate cv per biological replicate
cv_fun <- function(x) {
  # falls mean == 0 -> NA vermeiden
  m <- mean(x, na.rm = TRUE)
  if (is.na(m) || m == 0) return(NA_real_)
  sd(x, na.rm = TRUE) / m
}
# define pairs
pairs <- list(c(1,2), c(3,4), c(5,6), c(7,8))
# check if columns exist
pairs <- Filter(function(cols) all(cols <= ncol(df_mean)), pairs)

# rowwise cv
cv_rowwise <- sapply(pairs, function(cols) {
  apply(df_mean[, cols, drop = FALSE], 1, function(x) cv_fun(x))
})
cv_rowwise <- as.data.frame(cv_rowwise)
# name columns
colnames(cv_rowwise) <- c("Control", "Etoposide", "RE", "Resveratrol")[seq_along(pairs)]

# long format for plotting
cv_long <- pivot_longer(as.data.frame(cv_rowwise),
                        cols = everything(),
                        names_to = "Treatment",
                        values_to = "cv")

ggplot(cv_long, aes(x = Treatment, y = cv)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    x = "Spaltenpaar",
    y = "CV"
    )


colMeans(cv_rowwise*100, na.rm = TRUE)
apply(cv_rowwise*100, 2, median, na.rm = TRUE)

result_list <- lapply(colnames(cv_rowwise), function(col) {
  
  cv_values <- cv_rowwise[[col]]
  
  data.frame(
    Group = col,
    Median_CV = median(cv_values, na.rm = TRUE),
    IQR_CV = IQR(cv_values, na.rm = TRUE),
    Range_CV = max(cv_values, na.rm = TRUE) - min(cv_values, na.rm = TRUE),
    CV_lt_10 = mean(cv_values < 0.1, na.rm = TRUE) * 100,
    CV_lt_20 = mean(cv_values < 0.2, na.rm = TRUE) * 100,
    CV_gt_30 = mean(cv_values > 0.3, na.rm = TRUE) * 100
  )
})

summary_df <- do.call(rbind, result_list)
summary_df
```



### PCA

```{r}

df.LFQ <- df.prot[,c(260:262, 266:271,
                     272:277, 278:283,
                     296:301, 302:307,
                     284:289, 290:295)]

df.LFQ <- log2(df.LFQ)
fn <- colnames(df.LFQ)

# shorten filenames
fn_short <- sub("^LFQ\\.intensity\\.", "", fn)
fn_short <- sub("^.*CR_", "", fn_short)
fn_short <- sub("\\.raw\\.thermo\\.raw$", "", fn_short)

# rename them
colnames(df.LFQ) <- fn_short



mat <- df.LFQ[,c(1:3,
                 10:15,
                 22:27,
                 34:39)]

colnames(mat)

mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)



plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",        
  text = ~Filename,
  textposition = "top center",  
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```

```{r}
mat <- df.LFQ[,-c(1:3,
                 10:15,
                 22:27,
                 34:39)]

colnames(mat)


mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)



plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",        
  text = ~Filename,
  textposition = "top center",  
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```




# l-428

## Control

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(500:523)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:24
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Etoposide

```{r}
############################################################
# E
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(524:547)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:24
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## RE

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(572:586, 588:595)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:23
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## R

```{r}
############################################################
# R
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(548:571)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = FALSE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:24
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Compare All

```{r}
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

df.prot[df.prot == 0] <- NA

```

### CV calculation

```{r}
df.LFQ <- df.prot[,c(500:523,
                     524:547,
                     572:586, 588:595,
                     548:571)]

n <- ncol(df.LFQ)

block_sizes <- c(9, 6, 9, 
                 9, 6, 9,
                 9, 6, 8,
                 9, 6, 9)

if (sum(block_sizes) != n) stop("block sizes do not match column number")

ends <- cumsum(block_sizes)
starts <- c(1, head(ends, -1) + 1)

starts
ends

# generate df mean
df_mean <- as.data.frame(
  sapply(seq_along(starts),
         function(i) rowMeans(df.LFQ[, starts[i]:ends[i], drop = FALSE], na.rm = TRUE))
)

# col names
colnames(df_mean) <- paste0("block", seq_along(starts))

# calculate cv per biological replicate
cv_fun <- function(x) {
  # falls mean == 0 -> NA vermeiden
  m <- mean(x, na.rm = TRUE)
  if (is.na(m) || m == 0) return(NA_real_)
  sd(x, na.rm = TRUE) / m
}
# define pairs
pairs <- list(c(1,2,3), c(4,5,6), c(7,8,9), c(10,11,12))
# check if columns exist
pairs <- Filter(function(cols) all(cols <= ncol(df_mean)), pairs)

# rowwise cv
cv_rowwise <- sapply(pairs, function(cols) {
  apply(df_mean[, cols, drop = FALSE], 1, function(x) cv_fun(x))
})
cv_rowwise <- as.data.frame(cv_rowwise)
# name columns
colnames(cv_rowwise) <- c("Control", "Etoposide", "RE", "Resveratrol")[seq_along(pairs)]

# long format for plotting
cv_long <- pivot_longer(as.data.frame(cv_rowwise),
                        cols = everything(),
                        names_to = "Treatment",
                        values_to = "cv")

ggplot(cv_long, aes(x = Treatment, y = cv)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    x = "Spaltenpaar",
    y = "CV"
    )


colMeans(cv_rowwise*100, na.rm = TRUE)
apply(cv_rowwise*100, 2, median, na.rm = TRUE)


result_list <- lapply(colnames(cv_rowwise), function(col) {
  
  cv_values <- cv_rowwise[[col]]
  
  data.frame(
    Group = col,
    Median_CV = median(cv_values, na.rm = TRUE),
    IQR_CV = IQR(cv_values, na.rm = TRUE),
    Range_CV = max(cv_values, na.rm = TRUE) - min(cv_values, na.rm = TRUE),
    CV_lt_10 = mean(cv_values < 0.1, na.rm = TRUE) * 100,
    CV_lt_20 = mean(cv_values < 0.2, na.rm = TRUE) * 100,
    CV_gt_30 = mean(cv_values > 0.3, na.rm = TRUE) * 100
  )
})

summary_df <- do.call(rbind, result_list)
summary_df
```

### PCA 


```{r}

df.LFQ <- df.prot[,c(500:523,
                     524:547,
                     572:586, 588:595,
                     548:571)]


df.LFQ <- log2(df.LFQ)
fn <- colnames(df.LFQ)

# shorten filenames
fn_short <- sub("^LFQ\\.intensity\\.", "", fn)
fn_short <- sub("^.*HL_428_", "", fn_short)
fn_short <- sub("\\.raw\\.thermo\\.raw$", "", fn_short)

# rename them
colnames(df.LFQ) <- fn_short


mat <- df.LFQ[,c(1:9,
                 25:33,
                 49:57,
                 72:80)]
colnames(mat)

mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)


plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",        
  text = ~Filename,
  textposition = "top center", 
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```

```{r}
mat <- df.LFQ[,c(10:15,
                 34:39,
                 58:63,
                 81:86)]
colnames(mat)
mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)


plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",        
  text = ~Filename,
  textposition = "top center", 
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```

```{r}



mat <- df.LFQ[,c(16:24,
                 40:48,
                 64:71,
                 87:95)]

colnames(mat)

mat <- as.data.frame(lapply(mat, as.numeric))
mat <- mat[rowSums(!is.na(mat)) > 0, , drop = FALSE]
mat_cc <- mat[complete.cases(mat), , drop = FALSE]
v <- apply(mat_cc, 1, var)
mat_cc <- mat_cc[v > 0, , drop = FALSE]
pca_res <- prcomp(t(mat_cc), center = TRUE, scale. = FALSE)

# plot it
scores <- as.data.frame(pca_res$x)
scores$Filename <- rownames(scores)


plot_ly(
  scores,
  x = ~PC1, 
  y = ~PC2,
  type = "scatter",
  mode = "markers+text",        
  text = ~Filename,
  textposition = "top center", 
  marker = list(size = 10)
) %>%
  plotly::layout(
    xaxis = list(title = paste0("PC1 (", 
                                round(summary(pca_res)$importance[2,1] * 100, 1), "%)")),
    yaxis = list(title = paste0("PC2 (", 
                                round(summary(pca_res)$importance[2,2] * 100, 1), "%)"))
  )





```
