---
output:
  html_document:
    df_print: paged
    fig.align: center
    self_contained: yes 
    fig.height: 4
    fig.width: 8
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
    number_sections: yes
    code_folding: hide
title: "Reproducibility rev001"
author: "David Lilek"
date: "`r format(Sys.time(), '%d %B %Y, %X')`"
editor_options: 
  markdown: 
    wrap: 72
---

# Load Libraries & Plot function & Colors

```{r message=FALSE, warning=FALSE}
# load libaries
library(GGally)
library(corrplot)
library(ggplot2)
library(pheatmap)
library(tidyverse)
library(dplyr)
library(UpSetR)
library(knitr)
library(reshape2)
library(eulerr)
library(UniprotR)
library(stringr)
library(ggfittext)
```

# l-540 living

## Control

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(320:334)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```
```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```



## Etoposide

```{r}
############################################################
# E
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]


# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(335:349)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```



## RE

```{r}
############################################################
# RE
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(365:379)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```


## R

```{r}
############################################################
# R
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_le_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]


# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(350:364)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))

```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```



# l-540 cryo

## Control

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]


# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(260:262, 266:271)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:3
B <- 4:9

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Etoposide

```{r}
############################################################
# E
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(272:277, 278:283)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:7
B <- 8:12

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## RE

```{r}
############################################################
# RE
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(296:301, 302:307)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:6
B <- 7:12

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```
```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## R

```{r}
############################################################
# R
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl540_cryo_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(284:289, 290:295)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:6
B <- 7:12

rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
plot(
  rep_means$mean_A,
  rep_means$mean_B,
  xlab = "Mean log2 LFQ (Group A)",
  ylab = "Mean log2 LFQ (Group B)",
  main = "Replicate comparison",
  pch = 16, col = rgb(0, 0, 0, 0.4)
)

abline(0, 1, col = "red", lwd = 2)


data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```
```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

# l-428

## Control

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(500:523)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:24
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## Etoposide

```{r}
############################################################
# E
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(524:547)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:24
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## RE

```{r}
############################################################
# C
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(572:586, 588:595)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```

```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:23
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```

## R

```{r}
############################################################
# R
############################################################

# Load MaxQuant output
df <- read.csv("./results_maxquant/hl428_all_galaxy.txt", sep = "\t")
df.prot <- df

# Remove reverse/contaminant/site-only identifications
df.prot <- df.prot[!df.prot$Reverse == "+", ]
df.prot <- df.prot[!df.prot$Potential.contaminant == "+", ]
df.prot <- df.prot[!df.prot$Only.identified.by.site == "+", ]

# Extract LFQ intensity columns (as provided)
df.LFQ <- df.prot[, c(548:571)]
df.LFQ[df.LFQ == 0] <- NA; print(colnames(df.LFQ))
```


```{r}
# ensure numeric matrix
df.LFQ.log2 <- as.data.frame(lapply(df.LFQ, as.numeric))

# log2 transformation
df.LFQ.log2 <- log2(df.LFQ.log2)

# replace ±Inf with NA
df.LFQ.log2[is.infinite(as.matrix(df.LFQ.log2))] <- NA

# remove rows containing only NA values
df.LFQ.log2 <- df.LFQ.log2[rowSums(is.na(df.LFQ.log2)) < ncol(df.LFQ.log2), ]

# shorten col names
colnames(df.LFQ.log2) <- gsub(
  "\\.raw\\.thermo\\.raw$", "",
  gsub("^HL_", "",
       gsub("^LFQ\\.intensity\\.", "", colnames(df.LFQ.log2)))
)

```


```{r}
paletteLength <- 20

myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(df.LFQ.log2)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(df.LFQ.log2),
         show_rownames = FALSE)
```


```{r}
# -------------------------
# 1) Prepare data (NO imputation)
# -------------------------

# proteins (rows) x samples (columns)
X <- as.matrix(df.LFQ.log2)

# keep only proteins without any NA values
X.cc <- X[complete.cases(X), , drop = FALSE]

# transpose: samples as rows, proteins as variables
X_t <- t(X.cc)

# -------------------------
# 2) PCA
# -------------------------
pca <- prcomp(X_t, center = TRUE, scale. = TRUE)

# variance explained
pvar <- (pca$sdev^2) / sum(pca$sdev^2)

# -------------------------
# 3) Score plot (each sample different color)
# -------------------------
scores <- data.frame(
  Sample = rownames(pca$x),
  PC1 = pca$x[, 1],
  PC2 = pca$x[, 2]
)

ggplot(scores, aes(x = PC1, y = PC2, color = Sample)) +
  geom_point(size = 3) +
  labs(
    x = paste0("PC1 (", round(100 * pvar[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * pvar[2], 1), "%)"),
    title = "PCA of log2(LFQ intensities)"
  ) +
  theme_classic() +
  theme(legend.position = "right")

```


```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(df.LFQ.log2, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)

```
```{r}
par(mfrow = c(4, 4), mar = c(3, 3, 2, 1))  # layout anpassen

for (i in seq_len(ncol(df.LFQ.log2))) {
  x <- df.LFQ.log2[, i]
  x <- x[is.finite(x)]
  
  hist(
    x,
    breaks = 50,
    main = colnames(df.LFQ.log2)[i],
    xlab = "log2(LFQ)",
    col = "grey",
    border = "white"
  )
}

par(mfrow = c(1, 1))
```


### compare replicates

```{r}
# define replicate groups by column indices
A <- 1:9
B <- 10:15
C <- 16:24
rep_means <- data.frame(
  mean_A = rowMeans(df.LFQ.log2[, A], na.rm = TRUE),
  mean_B = rowMeans(df.LFQ.log2[, B], na.rm = TRUE),
  mean_C = rowMeans(df.LFQ.log2[, C], na.rm = TRUE)
)

```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data <- rep_means

# correlation scatter
ggpairs(data,
        title = "Pairwise correlation of samples",
        upper = list(continuous = wrap("cor", size = 4)),
        diag = list(continuous = "densityDiag"),
        lower = list(continuous = "smooth"))
```

```{r}
paletteLength <- 20
myColor <- colorRampPalette(c("navy", "white",
                                  "firebrick3"))(paletteLength)
pheatmap(as.matrix(na.omit(rep_means)),
         color = myColor,
         treeheight_row = 10,
         angle_col ="45",
         treeheight_col = 10,
         legend = TRUE,
         labels_col = colnames(rep_means),
         show_rownames = FALSE)
```

```{r}
# compute correlation between samples (columns)
cor_matrix <- cor(rep_means, use = "pairwise.complete.obs", method = "pearson")

# draw heatmap of correlations
pheatmap(cor_matrix,
         color = colorRampPalette(c("navy", "white", "firebrick3"))(50),
         main = "Sample Correlation Matrix",
         treeheight_row = 15,
         treeheight_col = 15,
         fontsize_col = 8,
         fontsize_row = 8)
```


```{r}
# remove na values and sort it
data <- na.omit(rep_means)
# "long format" for ggplot
ta_long <- data %>%
  mutate(Protein = rownames(.)) %>%
  pivot_longer(cols = starts_with("mean_"),
               names_to = "Group",
               values_to = "log2_Intensity") %>%
  group_by(Group) %>%
  arrange(desc(log2_Intensity), .by_group = TRUE) %>%
  mutate(Rank = row_number()) %>%
  ungroup()

# Plot
ggplot(ta_long, aes(x = Rank, y = log2_Intensity, color = Group)) +
  geom_point(size = 2) +
  geom_line() +
  labs(title = "Protein Rank Plot",
       x = "Protein abundance rank (log2)",
       y = "log2 Intensity") +
  theme_minimal()

```
